---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document: 
    keep_md: true
    author: "Saumya Bhardwaj"
---

*Saumya Bhardwaj*

## Loading and preprocessing the data
Before load the data, we will load the ```dplyr``` and ```ggplot2```  libraries. 
```{r message=FALSE}
library(dplyr)
library(ggplot2)
```

Then, unzip and load the data. Then convert the ```date``` collumn into Date format. And then, group the dataframe by date, and summarize to get the total, mean, median, minimum and maximum steps per date.

```{r}
unzip("activity.zip",exdir="./data")
data <- read.csv("./data/activity.csv")
data$date <- as.Date(data$date)
summary_data <- data %>% group_by(date) %>%
  summarise(total=sum(steps,na.rm=T),mean=mean(steps,na.rm = T),
            median=median(steps,na.rm = T),
            maximum=max(steps),minimum=min(steps),.groups = 'drop')
```

Then we will look at the first five rows in this new dataset
```{r}
head(summary_data)
```
## What is mean total number of steps taken per day?

The histogram of total steps per day is given by the next code
```{r}
with(summary_data,hist(total,breaks=10,main="Histogram of Total Steps per day",
                       col="blue",border="orange"))
```

The mean and median total can be easily calculated by
```{r}
mean_total <- mean(summary_data$total,na.rm=T)
median_total <- median(summary_data$total,na.rm=T)
mean_total
median_total
```

So, the mean total is `r mean_total` and the median total is `r median_total`.  

## What is the average daily activity pattern?

Now we are going to plot a time series of the average steps per interval. But first, we need to summarize our data grouped by interval. We will look to the first 5 lines of this new dataframe.

```{r}
interval_data <- data %>% group_by(interval) %>%
  summarise(total=sum(steps,na.rm=T),mean=mean(steps,na.rm = T),
            median=median(steps,na.rm = T),.groups = 'drop')
head(interval_data)            
```
To plot the time series needed, we are going to use the ```ggplot2``` library.

```{r}
qplot(interval,mean,data=interval_data,geom='line',
       xlab="Interval",main="Time Series of average steps per interval",
       ylab="Average",colour=I("red"))
```

The interval that have on average the maximum number of steps can be found with the next line of code

```{r}
interval_data[which.max(interval_data$total),][[1]]
```


## Imputing missing values

To determine the amount of NA's in the dataset, we can use the ```table``` function

```{r}
table(is.na(data$steps))
```

So, the total number of missing values is `r table(is.na(data$steps))[[2]]`. That represents `r round(table(is.na(data$steps))[[2]]/dim(data)[[1]]*100,2)` percent of the whole dataset, so we will fill the missing the data using the mean of the interval. For that, we will use the values already calculated in ```interval_data```. So, we will begin creating ```data_filled``` using the main dataset. Then we will search for the indices of the rows that have na's, then save the interval where it happens, and retrieve the result from the ```interval_data``` dataset.

```{r}
data_filled <- data
for (i in which(is.na(data_filled$steps))){
  missedInterval <- data_filled[i,3]
  data_filled[i,1] <- interval_data[[which(interval_data$interval==missedInterval),3]]
}
```
